A Real-Time Sonar Simulator for Automatic Target Recognition

### 2015-02-26 Simulator/article.tex
Changed simulator title
la2pdf commit - 2015:02:26	18:05:17	(Thu26.Feb)

% \usepackage{caption}
\usepackage{subcaption}
\usepackage{color}
\usepackage{calc}
\usepackage{fp}

@@

\title{A GPU Sonar Simulator for Automatic Target Recognition}

@@

Automatic target recognition (ATR) is an important component in autonomous underwater vehicles (AUVs), as it allows the vehicle to adapt its mission plan to e.g. revisit detected objects for closer examination. One way to classify these objects is by using template matching. The principle involved  is to isolate an image segment containing the object of interest, compare it with a set of template images of the relevant object classes, and assign it to the class of the template with the the best fit.

The challenge is to come up with accurate template images covering the actual configurations of object and seabed.

Estimate 2D template directly
  ->

>>>>

% \usepackage{caption}
% \usepackage{subcaption}
\usepackage{color}
\usepackage{calc}
\usepackage{fp}

@@

\title{A Real-Time Sonar Simulator for Automatic Target Recognition}

@@

One of the major shortcomings of autonomous underwater vehicles (AUVs) is their lack of curiosity. When encountering an interesting object, an AUV with a mounted sonar imaging system should be able to detect it, and perhaps move in for closer examination. Besides the apparent challenge of trying to detect something with a long list of unknown parameters, this must happen in real-time and be very power efficient.

% Automatic target recognition (ATR) is an important component in autonomous underwater vehicles (AUVs), as it allows the vehicle to adapt its mission plan to e.g. revisit detected objects for closer examination. One way to classify these objects is by using template matching. The principle involved  is to isolate an image segment containing the object of interest, compare it with a set of template images of the relevant object classes, and assign it to the class of the template with the the best fit.

The detection process is well described in the field of automatic target recognition (ATR). The classical approach involves isolating an object of interest, extracting parameters from it and feeding those into a classifier. However, it is hard to find parameters that describe the objects uniquely, in particular due to the ambiguities arising from interpreting a 3D world from its 2D projection. Another approach is to use template matching, where the image segment is compared to a set of template images covering the relevant object classes, and assigned to the class of the template yielding the the best fit. The challenge here is to come up with accurate template images covering the actual configurations of object and seabed, especially if a static template library is used, which will lead to a computational complexity that is exponential with the number of model parameters. This limits its use to only a few parameters being coarsely sampled, ultimately yielding inaccurate results~\cite{Midelfart2010}.

We have developed a synthetic aperture sonar (SAS) simulator that avoids this problem by running sufficiently fast to create templates adapted to the actual scene as part of the classification process. This is achieved with the aid of the massive computing power in graphics processing units (GPUs) and optimized software libraries for scene rendering. The simulator loads a 3D model of the seafloor and an object class model into OpenGL, where emitted sound waves are modeled with a light tube extending all the way down the propagation path of the AUV. We assume rough, isotropic surfaces reflecting sound energy equally in all directions. This can be modeled with a Lambertian scattering model~\cite{Blake1993,Bell1995}, where the intensity of the backscattered sound solely depends on the incidence angle of the transmitted signal onto the model surface. When rendering OpenGL is set up to produce an optical 2D intensity image and depth map that reveals the distance from each pixel to the propagation axis. For maximum flexibility this data is finally combined with OpenCL to produce the image templates.
